SunPy - Python for Solar Physics

### 2014-08-25 3-2-HEK.tex
created code text files

The Sun is an active star and exhibits a wide range of transient phenomena 
(e.g., flares, radio bursts) on many different time-scales, length-scales, and 
wavelengths. Observations and metadata concerning these phenomena are collected 
in the Heliophysics Event Knowledgebase (HEK) \citep{hek2012}.  Entries are generated both by 
automated algorithms and human observers.  Some of the information in the HEK 
reproduces feature and event data from elsewhere (for example, the \textit{GOES} flare catalogue),
and some is generated by the Solar Dynamics Observatory Feature Finding Team 
\citep{martens2012}.  The advantage of the HEK is it 
provides an homogeneous and well-described interface to a large amount of 
feature and event information of interest to the solar physics community. SunPy 
accesses this information through the \texttt{hek} module.  The \texttt{hek} module makes use of the 
HEK API\footnote{For more information see \url{http://vso.stanford.edu/hekwiki/ApplicationProgrammingInterface}}.

Simple HEK queries consist of start time, an end time, and an event type 
(Listing~\ref{code:hek:simple}). Event types are specified as upper case, 
two letter strings, and these strings are 
identical to the two letter abbreviations found at the HEK website
(\url{http://www.lmsal.com/hek/VOEvent_Spec.html}).

@@

The module \texttt{hek.attrs} contains attributes of the HEK that can be used to
construct HEK queries.  For example, a flare is an attribute of the HEK,
 and so 
\texttt{hek.attrs.EventType("FL")} in Listing~\ref{code:hek:simple}
can be alternatively replaced with \texttt{hek.attrs.FL}.

HEK attributes differ from VSO attributes (Section \ref{ssec:vso}) in that many 
of them are wrappers that conveniently expose comparisons by overloading Python 
operators. This allows filtering of the HEK entries by the properties of the 
event. As was mentioned above, the HEK stores feature/event metadata obtained 
in different ways, known generally as feature recognition methods (FRMs). 
Example in Listing~\ref{code:hek:frm} filters the results of the previous 
result to return only those events that have the FRM `SSW Latest Events'.  Multiple comparisons can be made by including more comma-separated
conditions on the attributes in the call to the HEK query method.

@@

HEK comparisons can be combined using Python's logical operators (e.g., \texttt{and}
and \texttt{or}). This makes complex queries easy to create: Listing \ref{code:hek:or} 
returns flares west of 50 arcseconds or those that have a peak flux above 
1000.0 (the units of the flux are FRM-dependent and are described at the HEK 
website).
\begin{listing}[H]
\begin{minted}[bgcolor=bg]{python}
>>> result = client.query(hek.attrs.Time(tstart, tend), 
...                       hek.attrs.EventType("FL"),
...                       (hek.attrs.Event.Coord1>50) 
...                       or (hek.attrs.FL.PeakFlux>1000.0))
\end{minted}
\caption{HEK query using the logical \texttt{or} operator.}
\label{code:hek:or}
\end{listing}
All FRMs report the required feature attributes, but the optional attributes 
are FRM dependent.  If a FRM does not have one of the optional attributes, 
\texttt{None} is returned by the \texttt{hek} module.

The ability to use comparison and logical operators on HEK attributes allows 
the construction of queries of almost arbitrary complexity. 
The results of a HEK query can be used to download the 
corresponding data from the VSO using SunPy's \texttt{H2VClient}, as 
demonstrated in Listing~\ref{code:hek2vso}.
\begin{listing}[H]
\begin{minted}[bgcolor=bg]{python}
>>> from sunpy.net import hek2vso
>>> h2v = hek2vso.H2VClient()
>>> vso_results = h2v.translate_and_query(result[0])
>>> h2v.vso_client.get(vso_results[0]).wait()
\end{minted}
\caption{Code snippet continuing from Listing~\ref{code:hek:or} showing the 
query and download of data from the first HEK result from the VSO.}
\label{code:hek2vso}
\end{listing}

>>>>

The Sun is an active star and exhibits a wide range of transient phenomena 
(e.g., flares, radio bursts, coronal mass ejections) at many different time-scales, 
length-scales, and 
wavelengths. Observations and metadata concerning these phenomena are collected 
in the Heliophysics Event Knowledgebase (HEK) \citep{hek2012}.  Entries are generated both by 
automated algorithms and human observers.  Some of the information in the HEK 
reproduces feature and event data from elsewhere (for example, the \textit{GOES} flare catalogue),
and some is generated by the Solar Dynamics Observatory Feature Finding Team 
\citep{martens2012}.  A key feature of the HEK is that it
provides an homogeneous and well-described interface to a large amount of 
feature and event information of interest to the solar physics community. SunPy 
accesses this information through the \texttt{hek} module.  The \texttt{hek} module makes use of the 
HEK public API\footnote{For more information see \url{http://vso.stanford.edu/hekwiki/ApplicationProgrammingInterface}}.

Simple HEK queries consist of start time, an end time, and an event type 
(see Listing~\ref{code:hek:simple}). Event types are specified as upper case, 
two letter strings, and these strings are 
identical to the two letter abbreviations defined by HEK 
(see \url{http://www.lmsal.com/hek/VOEvent_Spec.html}). Users can see a
complete list and description of these abbreviations by looking at the documentation
for \texttt{hek.attrs.EventType}.

@@

Short-cuts are also provided for some often-used event types. For example, 
the flare attribute can be declared as either 
\texttt{hek.attrs.EventType("FL")} or as \texttt{hek.attrs.FL}.

HEK attributes differ from VSO attributes (Section \ref{ssec:vso}) in that many 
of them are wrappers that conveniently expose comparisons by overloading Python 
operators. This allows filtering of the HEK entries by the properties of the 
event. As was mentioned above, the HEK stores feature and event metadata obtained 
in different ways, known generally as feature recognition methods (FRMs). 
Example in Listing~\ref{code:hek:frm} filters the results of the previous 
result to return only those events that have the FRM `SSW Latest Events'.  
Multiple comparisons can be made by including more comma-separated
conditions on the attributes in the call to the HEK query method.

@@

HEK comparisons can be combined using Python's logical operators (e.g., \texttt{and}
and \texttt{or}). The ability to use comparison and logical operators on HEK attributes allows 
the construction of queries of almost arbitrary complexity.
For the query in Listing \ref{code:hek:or} returns
returns flares with helioprojective x coordinates west of 50 arcseconds or 
those that have a peak flux above 1000.0 in units defined by the FRM.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{python}
>>> result = client.query(hek.attrs.Time(tstart, tend), 
...                       hek.attrs.EventType("FL"),
...                       (hek.attrs.Event.Coord1>50) 
...                       or (hek.attrs.FL.PeakFlux>1000.0))
\end{minted}
\caption{HEK query using the logical \texttt{or} operator.}
\label{code:hek:or}
\end{listing}
All FRMs report the required feature attributes, but the optional attributes 
are FRM dependent.  If a FRM does not have one of the optional attributes, 
\texttt{None} is returned by the \texttt{hek} module.

Frequently after users have found events of interest the next step is to 
download observational data. The \texttt{H2VClient} module makes this
much easier by providing a translation layer between HEK query results
and VSO queries. This capability is demonstrated in Listing~\ref{code:hek2vso}.
\begin{listing}[H]
\begin{minted}[bgcolor=bg]{python}
>>> from sunpy.net import hek2vso
>>> h2v = hek2vso.H2VClient()
>>> vso_results = h2v.translate_and_query(result[0])
>>> h2v.vso_client.get(vso_results[0]).wait()
\end{minted}
\caption{Code snippet continuing from Listing~\ref{code:hek:or} showing the 
query and download of data from the first HEK result from the VSO.}
\label{code:hek2vso}
\end{listing}

### 2014-08-25 4-Util.tex
removed spaces
created code text files

The \texttt{sun.constants} module provides a number of solar-related 
constants in order to provide consistency in the calculations of derived solar 
values  within the SunPy code base, but also to the user. Every solar 
constant is provided as a \texttt{Constant} object as defined by Astropy. Each 
\texttt{Constant} object defines a \texttt{Quantity}, a number associated with a unit, along with 
the constant's provenance (i.e., reference) and its uncertainty. Using 
Astropy's \texttt{Quantity} objects, any solar constant can easily be converted between 
different units, including between the SI or cgs unit systems, as can be seen in Listing~\ref{code:constants_code}.
As these objects inherit from 
NumPy's \texttt{ndarray}, they work well with standard representations of numbers.
For convenience, a number of shortcuts to frequently used constants are provided 
directly when importing the module. A larger list of constants can be 
accessed through an interface modelled on that provided by the SciPy constants 
package and is available as a dictionary called \texttt{physical\_constants}. 
To view them all quickly, a \texttt{print\_all()} function is available.

>>>>

The \texttt{sun.constants} module provides a number of solar-related 
constants in order to provide consistency in the calculations of derived solar 
values within SunPy, but also to the user. Every solar 
constant is provided as a \texttt{Constant} object as defined by the Astropy \texttt{units} package. Each 
\texttt{Constant} object defines a \texttt{Quantity}, a number associated with a unit, along with 
the constant's provenance (i.e., reference) and its uncertainty. Using 
Astropy's \texttt{Quantity} objects, any solar constant can easily be converted between 
different units, including between the SI or cgs unit systems, as can be seen in Listing~\ref{code:constants_code}.
As these objects inherit from 
NumPy's \texttt{ndarray}, they work well with standard representations of numbers.
For convenience, a number of shortcuts to frequently used constants are provided 
directly when importing the module. A larger list of constants can be 
accessed through an interface modeled on that provided by the SciPy constants 
package and is available as a dictionary called \texttt{physical\_constants}. 
To view them all quickly, a \texttt{print\_all()} function is available.
